When are Fuzzy Extractors Possible?

### 2015-02-10 trunk/info-bounds/FuzzyExtPossibility.tex
minor edits from reviewer comments

\begin{abstract}
Fuzzy extractors (Dodis et al., Eurocrypt 2004) convert repeated noisy readings of a high-entropy secret into the same uniformly distributed key. A minimum condition for the security of the key is the hardness of guessing a value that is similar to the secret, because the fuzzy extractor converts such a guess to the key.
We define \emph{fuzzy min-entropy} to quantify this property of a noisy source of secrets.  Fuzzy min-entropy measures the success of the adversary when provided with \emph{only} the functionality of the fuzzy extractor and thus the \emph{ideal} security possible from a noisy distribution.

High fuzzy min-entropy is necessary for the existence of a fuzzy extractor; moreover, there is evidence that it may be sufficient when only computational security is required. Nevertheless, information-theoretic fuzzy extractors are not known for many practically relevant sources of high fuzzy min-entropy. The challenge is to design information-theoretic fuzzy extractors whose security (almost) matches the fuzzy min-entropy of a distribution.
In this work, we ask whether this challenge is inherent.
 We provide two main technical results.
\begin{itemize}
\item \textbf{Sufficiency of fuzzy min-entropy:}  We construct a fuzzy extractor for each physical source with fuzzy min-entropy.  This result justifies fuzzy min-entropy as the right measure of the quality of a noisy distribution for key derivation.  Previous results considered the entropy and error tolerance of a separately and had not secured many sources with fuzzy min-entropy.  The constructed fuzzy extractor must \emph{precisely know} the distribution of the noisy source.
\item \textbf{Cost of distributional uncertainty:}  High entropy sources can never be fully observed and must be modeled.  An adversary may have a better model than the fuzzy extractor designer.  To compensate, fuzzy extractors are typically designed to work for families of physical sources. We show that this uncertainty is an impediment to security by building a family of high fuzzy min-entropy sources for which no fuzzy extractor can exist.  This result emphasizes the importance of high fidelity models of high entropy sources.
\end{itemize}
We provide similar but stronger results for secure sketches, whose goal is not to derive a consistent key, but to recover a consistent reading of the secret.  Our results are information-theoretic, but we discuss implications for the computational setting.

@@

Fuzzy min-entropy measures the quality of a distribution when provided oracle access to the functionality of a fuzzy extractor.  It can be thought of as measuring the ideal security of a noisy source of entropy when used for key derivation.  This is similar to how security is measured for program obfuscation~\cite{barak2001possibility}, we discuss the connection to program obfuscation in \secref{sec:related settings}.  Superlogarithmic fuzzy min-entropy is \emph{necessary} for nontrivial key extraction~(\propref{prop:fuzz necessary} formalizes the above intuition).

@@

\paragraph{The Challenge of Precisely Knowing a High Entropy Distribution}
Assuming \emph{precise knowledge} of a distribution $W$ may be unrealistic. Indeed, high-entropy distributions can never be fully observed directly and must therefore be modeled. It is imprudent to assume that the designer's model of a distribution is completely accurate---the adversary, with greater resources, would likely be able to build a better model. Therefore, fuzzy extractor designs cannot usually be tailored to one  particular source. Existing designs work for a family of sources (for example, all sources of min-entropy at least $m$ with at most $t$ errors). Thus, the design is fixed with a partially known distribution, and the adversary may therefore know more about the distribution than the designer of the fuzzy extractor.  We call this setting \emph{distributional uncertainty}.
We ask whether fuzzy min-entropy is sufficient in each of these settings.

\paragraph{The Cost of Distributional Uncertainty}
When the distribution is $W$ is uncertain, our results are negative.   In this setting, $W$ is only known to be in family of distributions $\mathcal{W}_Z$.
We construct a family $\mathcal{W}_Z$ where not even a 2-bit fuzzy extractor can be secure for most distributions in  $\mathcal{W}_Z$.  We emphasize that each distribution $W_z\in \mathcal{W}_Z$ has superlogarithmic fuzzy min-entropy---in fact, $\Hfuzz(W_z)=\Hoo(W_z)$, because all points in $W_z$ are distance at least $t$ apart. Our proof relies on high dimensionality of each $W_z$ and on perfect correctness of the $\rep$ procedure.
This result shows that uncertainty in the distribution $W$ is devastating to building a fuzzy extractor.  In particular, our positive results (\corref{cor:extension to fuzz ext}) show it is possible to build a fuzzy extractor for each $W_z$, but no construction can simultaneously secure the whole family.  Our result motivates further research into high fidelity descriptions of noisy physical sources.

\begin{table}[h]
\begin{center}
\begin{tabular}{c  l l }
 & Precise Knowledge & Distributional Uncertainty\\
\hline
Fuzzy Extractor & Yes~(\corref{cor:extension to fuzz ext}) & No~(\thref{thm:imposs fuzz ext})\\
\hline
Secure Sketch & Yes~(\corref{cor:extension to fuzz ext}) & No~(\thref{thm:imposs sketch})
\end{tabular}
\end{center}
\caption{Is fuzzy min-entropy sufficient to extract a superlogarithmic length key?  Results are information-theoretic.}
\label{tab:main results}
\end{table}

@@

\bnote{compare to current constructions?}
\subsection{Our Techniques}

@@

Then we show how the adversary can exploit the knowledge of the affine line to reduce the uncertainty about $w$ (in the secure sketch case) or $\key$ (in the fuzzy extractor case). 
In the secure sketch case, $ss$ can be used to find fixed points of $\rec(\cdot, ss)$ which, by the correctness requirement of the sketch, must be separated by minimum distance $t$. This means there aren't too many of them, so few can lie on an average line, permitting the adversary to guess one easily.

@@

\subsection{Related Settings}
\label{sec:related settings}
\paragraph{Other settings with close readings:  $\Hfuzz$ is sufficient}
The security definition of fuzzy extractors and secure sketches can be weakened to protect only against computationally bounded adversaries~\cite{fuller2013computational}.   In this computational setting, a single fuzzy extractor (or secure sketch) can simultaneously secure all possible distributions by using virtual grey-box obfuscation for all circuits~\cite{BitanskyCKP14}. The construction places into $p$ the obfuscated program for testing proximity to $w$ and outputting the appropriate value if the test passes.  This construction is secure when the adversary can rarely learn \key with oracle access to the program functionality.  The set of distributions with fuzzy min-entropy are exactly those where an adversary learns \key with oracle access to the functionality with negligible probability.\footnote{If this construction is used for a secure sketch, $W$ will remain unpredictable conditioned on $p$, but will not have pseudoentropy (see  \secref{sec:feas comp sec sketch} for details).  Extending our negative result to the computational setting would rule out the existence of semantically-secure multilinear maps.}

@@

\paragraph{Correlated rather than close readings}
A different model for the problem of key derivation from noisy sources does not explicitly consider the distance between $w$ and $w'$, but rather views $w$ and $w'$ as samples of drawn from a correlated pair of random variables.   This model is considered in multiple works, including~\cite{wyner1975wire,DBLP:journals/tit/CsiszarK78,DBLP:journals/tit/AhlswedeC93,DBLP:journals/tit/Maurer93}; recent characterizations of when key derivation is possible in this model include \cite{DBLP:conf/asiacrypt/RennerW05} and \cite{DBLP:journals/corr/TyagiW14}.  We compare our positive results to these characterizations in \apref{sec:corr variables}.  To the best of our knowledge, prior results on correlated random variables are in the precise knowledge setting, we are unaware of works that consider implications of distributional uncertainty.

%The definition of $\Hfuzz$ can be extended to this model (by again thinking of the adversary as choosing the value $w'$ that is most likely to lead to a success).

@@

The {\em statistical distance} between random variables $X$ and $Y$ with the same domain is $\sd(X,Y) = \frac12 \sum_x |\Pr[X=x] - \Pr[Y=x]|$.
%For a distinguisher $D$ we write the \emph{computational distance} between $X$ and $Y$ as $\delta^D(X,Y) = \left| \expe[D(X)]-\expe[D(Y)]\right |$ (we extend it to a class of distinguishers $\mathcal{D}$ by taking the maximum over all distinguishers $D\in\mathcal{D}$).  We denote by $\mathcal{D}_{s}$ the class of randomized circuits which output a single bit and have size at most $s$.
For a metric space $(\mathcal{M}, \dis)$, the \emph{(closed) ball of radius $t$ around $w$} is the set of all points within radius $t$, that is, $B_t(w) = \{w'| \dis(w, w')\leq t\}$.  If the size of a ball in a metric space does not depend on $w$, we denote by $|B_t|$ the size of a ball of radius $t$.  We consider the Hamming metric over vectors in $\mathcal{Z}^\gamma$, defined via $\dis(w,w') = \{i | w_i \neq w'_i\}$ where $\mathcal{Z}$ is some alphabet.  For this metric, $|B_t| = \sum_{i=0}^t {\gamma \choose i} (|\mathcal{Z}|-1)^i $.  $U_\kappa$ denotes the uniformly  distributed random variable on $\{0,1\}^\kappa$.  Throughout this work, we consider a sequence of metric spaces $\mathcal{M}_n$ parameterized by $n$, but we write $\mathcal{M}$ for notational convenience. A \emph{negligible} function $\ngl(n)$ is one that decreases faster than any positive inverse polynomial as $n\rightarrow \infty$.

@@

\begin{corollary}
\label{cor:imposs comp sketch}
Let $n$ be a security parameter and let $\mathcal{M} = |\mathbb{F}|^\gamma$.  There exists a family of distributions $\mathcal{W}_Z$ over $\mathcal{M}$ such that for each element $W_z\in \mathcal{W}_Z$, $\Hfuzz(W_z)= \omega(\log n)$ and for any $(\mathcal{M}, \mathcal{W}_Z, \tilde{m}, t)$-HILL secure sketch $(\sketch, \rec)$ that is $(s_{sec}, \epsilon_{sec})$-hard and error $\delta$.  If 
$s_{sec}\ge t(|\rec| + \gamma \log |\mathbb{F}|)$, 
$t\ge 4$, and 
 $\epsilon_{sec} + t\delta < 1/16$,
 then $\tilde{m} <4$.
\end{corollary}

>>>>

\begin{abstract}
Fuzzy extractors (Dodis et al., Eurocrypt 2004) convert repeated noisy readings of a high-entropy secret into the same uniformly distributed key. A minimum condition for the security of the key is the hardness of guessing a value that is similar to the secret, because the fuzzy extractor converts such a guess to the key.
We define \emph{fuzzy min-entropy} to quantify this property of a noisy source of secrets.  Fuzzy min-entropy measures the success of the adversary when provided with \emph{only} the functionality of the fuzzy extractor, that is, the \emph{ideal} security possible from a noisy distribution.

High fuzzy min-entropy is necessary for the existence of a fuzzy extractor; moreover, there is evidence that it may be sufficient when only computational security is required. Nevertheless, information-theoretic fuzzy extractors are not known for many practically relevant sources of high fuzzy min-entropy. In this work, we ask whether this challenge is inherent.  We show the following:
\begin{itemize}
\item \textbf{Sufficiency of fuzzy min-entropy:}  We construct a fuzzy extractor for each physical source with fuzzy min-entropy.  This result justifies fuzzy min-entropy as the right measure of the quality of a noisy distribution for key derivation.  Previous results considered the entropy and error tolerance of a separately and had not secured many sources with fuzzy min-entropy.  
The constructed fuzzy extractor must \emph{precisely know} the distribution of the noisy source.
\item \textbf{Cost of distributional uncertainty:}  High entropy sources can never be fully observed and must be modeled.  An adversary may have a better model than the fuzzy extractor designer.  To compensate, fuzzy extractors are typically designed to work for families of physical sources. We build a family of high fuzzy min-entropy sources which no fuzzy extractor can secure.  This result emphasizes the importance of high fidelity models of high entropy sources.
\end{itemize}
We provide similar but stronger results for secure sketches, whose goal is not to derive a consistent key, but to recover a consistent reading of the secret.  Our results are information-theoretic, but we discuss implications for the computational setting.

@@

Fuzzy min-entropy measures the quality of a distribution when provided oracle access to the functionality of a fuzzy extractor.  Fuzzy min-entropy measures the ideal security of a noisy source of entropy when used for key derivation, similar to distributional notions of program obfuscation~\cite{barak2001possibility,DBLP:conf/stoc/DodisS05,DBLP:conf/tcc/DodisS05}---we discuss the connection in \secref{sec:related settings}.  Superlogarithmic fuzzy min-entropy is \emph{necessary} for nontrivial key extraction~(\propref{prop:fuzz necessary} formalizes the above intuition).

@@

\paragraph{The Challenge of Precisely Distributional Knowledge}
Assuming \emph{precise knowledge} of a distribution $W$ may be unrealistic. Indeed, high-entropy distributions can never be fully observed directly and must therefore be modeled. It is imprudent to assume that the designer's model of a distribution is completely accurate---the adversary, with greater resources, would likely be able to build a better model. Therefore, fuzzy extractor designs cannot usually be tailored to one  particular source. Existing designs work for a family of sources (for example, all sources of min-entropy at least $m$ with at most $t$ errors). Thus, the design is fixed with a partially known distribution, and the adversary may know more about the distribution than the designer of the fuzzy extractor.  We call this the \emph{distributional uncertainty} setting.

\paragraph{The Cost of Distributional Uncertainty}
When the distribution is $W$ is uncertain, our results are negative.   In this setting, $W$ is known to be an element of a family of distributions $\mathcal{W}_Z$.
We construct a family $\mathcal{W}_Z$ where not even a 2-bit fuzzy extractor can be secure for most distributions in  $\mathcal{W}_Z$.  We emphasize that each distribution $W_z\in \mathcal{W}_Z$ has superlogarithmic fuzzy min-entropy---in fact, $\Hfuzz(W_z)=\Hoo(W_z)$, because all points in $W_z$ are distance at least $t$ apart. Our proof relies on high dimensionality of each $W_z$ and on perfect correctness of the $\rep$ procedure.
This result shows that uncertainty in the distribution $W$ is devastating to building a fuzzy extractor.  In particular, our positive results (\corref{cor:extension to fuzz ext}) show it is possible to build a fuzzy extractor for each $W_z$, but no construction can simultaneously secure the whole family.  Our result motivates further research into high fidelity descriptions of noisy sources.

\begin{table}[h]
\begin{center}
\begin{tabular}{c  l l }
 & Precise Knowledge & Distributional Uncertainty\\
\hline
Fuzzy Extractor & Yes~(\corref{cor:extension to fuzz ext}) & No~(\thref{thm:imposs fuzz ext})\\
\hline
Secure Sketch & Yes~(\corref{cor:extension to fuzz ext}) & No~(\thref{thm:imposs sketch})
\end{tabular}
\end{center}
\caption{Is fuzzy min-entropy sufficient to extract an information-theoretic superlogarithmic length key?}
\label{tab:main results}
\end{table}

@@

\paragraph{Interpreting our Results}
Our results show that fuzzy min-entropy is sufficient for noisy key derivation if the distribution is precisely known but insufficient if the distribution is only known to come from a family.  A similar situation occurs in the (noiseless) randomness extraction literature~\cite{nisan1993randomness}.  Deterministic randomness extractors are possible if the distribution is precisely known, but impossible for all high min-entropy distributions~\cite{santha1986generating}.  This impossibility can be overcome by the use of a short seed of uniform randomness.  
We must find an analogous relaxation to derive keys from noisy sources.  One natural relaxation is providing security against only  computationally bounded adversaries, see \secref{sec:related settings}.

\subsection{Our Techniques}

@@

An adversary can exploit the knowledge of the affine line to reduce the uncertainty about $w$ (in the secure sketch case) or $\key$ (in the fuzzy extractor case). 
In the secure sketch case, $ss$ can be used to find fixed points of $\rec(\cdot, ss)$ which, by the correctness requirement of the sketch, must be separated by minimum distance $t$. This means there aren't too many of them, so few can lie on an average line, permitting the adversary to guess one easily.

@@

\subsection{Related Settings}
\label{sec:related settings}
\paragraph{Other settings with close readings:  $\Hfuzz$ is sufficient}
The security definition of fuzzy extractors and secure sketches can be weakened to protect only against computationally bounded adversaries~\cite{fuller2013computational}.   In this computational setting, a single fuzzy extractor (or secure sketch) can simultaneously secure all possible distributions by using virtual grey-box obfuscation for all circuits~\cite{BitanskyCKP14}. The construction places into $p$ the obfuscated program for testing proximity to $w$ and outputting the appropriate value if the test passes.\footnote{If this construction is used for a secure sketch, $W$ will remain unpredictable conditioned on $p$, but will not have pseudoentropy (see  \secref{sec:feas comp sec sketch} for details).}  This construction is secure when the adversary can rarely learn \key with oracle access to the program functionality.  The set of distributions with fuzzy min-entropy are exactly those where an adversary learns \key with oracle access to the functionality with negligible probability. Thus, extending our negative result to the computational setting would rule out the existence of semantically-secure multilinear maps.

@@

\paragraph{Correlated rather than close readings}
A different model for the problem of key derivation from noisy sources does not explicitly consider the distance between $w$ and $w'$, but rather views $w$ and $w'$ as samples of drawn from a correlated pair of random variables.   This model is considered in multiple works, including~\cite{wyner1975wire,DBLP:journals/tit/CsiszarK78,DBLP:journals/tit/AhlswedeC93,DBLP:journals/tit/Maurer93}; recent characterizations of when key derivation is possible in this model include \cite{DBLP:conf/asiacrypt/RennerW05} and \cite{DBLP:journals/corr/TyagiW14}.  We compare our positive results to these characterizations in \apref{sec:corr variables}.  To the best of our knowledge, prior results on correlated random variables are in the precise knowledge setting, we are unaware of works that consider the cost of distributional uncertainty.

@@

The {\em statistical distance} between random variables $X$ and $Y$ with the same domain is $\sd(X,Y) = \frac12 \sum_x |\Pr[X=x] - \Pr[Y=x]|$.
%For a distinguisher $D$ we write the \emph{computational distance} between $X$ and $Y$ as $\delta^D(X,Y) = \left| \expe[D(X)]-\expe[D(Y)]\right |$ (we extend it to a class of distinguishers $\mathcal{D}$ by taking the maximum over all distinguishers $D\in\mathcal{D}$).  We denote by $\mathcal{D}_{s}$ the class of randomized circuits which output a single bit and have size at most $s$.
For a metric space $(\mathcal{M}, \dis)$, the \emph{(closed) ball of radius $t$ around $w$} is the set of all points within radius $t$, that is, $B_t(w) = \{w'| \dis(w, w')\leq t\}$.  If the size of a ball in a metric space does not depend on $w$, we denote by $|B_t|$ the size of a ball of radius $t$.  We consider the Hamming metric over vectors in $\mathcal{Z}^\gamma$, defined via $\dis(w,w') = |\{i | w_i \neq w'_i\}|$ where $\mathcal{Z}$ is some alphabet.  For this metric, $|B_t| = \sum_{i=0}^t {\gamma \choose i} (|\mathcal{Z}|-1)^i $.  $U_\kappa$ denotes the uniformly  distributed random variable on $\{0,1\}^\kappa$.  Throughout this work, we consider a sequence of metric spaces $\mathcal{M}_n$ parameterized by $n$, but we write $\mathcal{M}$ for notational convenience. A \emph{negligible} function $\ngl(n)$ is one that decreases faster than any positive inverse polynomial as $n\rightarrow \infty$.

@@

\begin{corollary}
\label{cor:imposs comp sketch}
Let $n$ be a security parameter and let $\mathcal{M} = |\mathbb{F}|^\gamma$.  There exists a family of distributions $\mathcal{W}_Z$ over $\mathcal{M}$ such that for each element $W_z\in \mathcal{W}_Z$, $\Hfuzz(W_z)= \omega(\log n)$ such that for any $(\mathcal{M}, \mathcal{W}_Z, \tilde{m}, t)$-HILL secure sketch $(\sketch, \rec)$ that is $(s_{sec}, \epsilon_{sec})$-hard and error $\delta$, if 
$s_{sec}\ge t(|\rec| + \gamma \log |\mathbb{F}|)$, 
$t\ge 4$, and 
 $\epsilon_{sec} + t\delta < 1/16$,
 then $\tilde{m} <4$.
\end{corollary}
